---
title: "Normalização Dataset 3"
output: html_notebook
---


```{r}

# Clear variables
rm(list=ls())
#Get path
path<-getwd(); path
# Set path
setwd(path)
```


```{r}
count_table <- read_delim("C:/Users/Acer/OneDrive/Ambiente de Trabalho/ICVS/Dataset 3/count_table.txt", 
    delim = "\t", escape_double = FALSE, 
    col_types = cols(...1 = col_skip()), 
    trim_ws = TRUE)

head(count_table)
```

## Expression Matrix

```{r}
expression.matrix = count_table[,-1]
```


## Data info

```{r}
design_tags <- read_delim("C:/Users/Acer/OneDrive/Ambiente de Trabalho/ICVS/Dataset 3/design_tags.tsv", 
    delim = "\t", escape_double = FALSE, 
    col_names = FALSE, trim_ws = TRUE)

head(design_tags)
```

## Rename columns

```{r}
data_info <- as.matrix(design_tags[, -1])
rownames(data_info) <- design_tags$X1

data_info = as.data.frame(data_info)
data_info$X2 = as.factor(data_info$X2)

levels(data_info$X2)
levels(data_info$X2) <- c("Control", "tgfBeta")

head(data_info)

```
## confirm that column names from expression matriz correspond to the row names of the data info

```{r}
all(rownames(data_info)==colnames(expression.matrix))
```


## Create a deseq object by importing the expression matrix and sample info

```{r}

dds <- DESeqDataSetFromMatrix(countData = expression.matrix,
                              colData = data_info,
                              design= ~ X2)

dds <- DESeq(dds)
resultsNames(dds) # lists the coefficients
res <- results(dds, name="X2_tgfBeta_vs_Control")
res
```
As noticeable, the row names are numbers instead of the Gene names, so next we will rename it

```{r}
names = c(count_table$Gene)

rownames(dds) <- names
head(dds)
```
It is better to filter the low count genes:

##Filtering

```{r}

keep = rowSums(counts(dds) >= 10) >= min(table(data_info$X2))
#>= min(table(data_info$X2))
# this condition implies that we have to see counts >= 10 at least in 3 samples because min(table(data_info$X2)) = 3
#see table(data_info$X2)

dds<- dds[keep,]

head(dds)
```

# Perform the statistical test(s) analysis to identify differentially expressed genes

## First Method
# standard analysis

```{r}
dds <- DESeq(dds)
res <- results(dds)


dds

res = as.data.frame(res)

head(res)

```

## Order DESeq results by padj

```{r}
res_ordered = res[order(res$pvalue),]
head(res_ordered)
```

```{r}
res05 <- results(dds, alpha=0.05)
summary(res05)
```

```{r}
sum(res05$padj < 0.05, na.rm=TRUE)
# LFC > 0 + LFC < 0 = 47 + 45 = 92
```
# CONTINUE to filter by padj values and Log2FoldChange values
res is the results from the dds 
```{r}

filtering <- res %>% filter(res$padj <0.05)

filtering <- filtering %>% filter(abs(filtering$log2FoldChange) >1)

filtering
```



# Variance stabelizing transformation
VST: variance-stabilizing transformation

```{r}
vsd = vst(dds, blind = FALSE)
```





